Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}

### 2023-06-08 paper/paper.tex
#22 rearrange schönholzer

In this paper I study the role of this process in the economic development of Europe. In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development.

\cite{levine2013, levine2021} argue that social evolution is driven by conflict. Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p. 436).Their answer to Why Europe and not China or India: "Ghengis Khan, the cannon, and the English. Ghengis Khan accomplished two things: he brought cannons to Europe, and he depopulated Mongolia. In China the removal of the outside threat of the Mongols led - as the theory asserts - to an extractive hegemony. In Europe the advent of the cannon together with the preying of the English on the French and the Spanish resulted - as the theory asserts - in an inclusive balance of power." (p. 439). strong outsiders but primitive siege technology [cheap defensibility of cities] in India led to extractive balance of power.

@@

@@

>>>>

In this paper I study the role of this process in the economic development of Europe. In particular, I estimate the effect of state power, the trait that conflict selects for, on urban development. If the effect is positive, then selection of "better" states through conflict-driven evolution may partially explain the ascendancy of Europe.

Past treatments of this topic have suffered from a lack of available historical data. \cite{diamond1997} and \cite{landes2006} each present narrative arguments for the economic benefits of state competition but stop short of formal empirical analysis. Such a study requires at the very least three things: 1) a detailed territorial history of European cities, 2) some city-level measure of development reaching far enough into the past, 3) some measure of state power, institutional quality, or whatever else one wishes to explain effects with.
These data limitations are starting to be overcome. Recently \cite{schoenholzer2022} have investigated how switching between states affects the development of cities. They find that switching states affects population negatively in the short run but positively in the long run. Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction that is compensated by the long-run benefit of becoming part of a "better" state.
However, they do not directly estimate the effect of anything other than switching to different states.

I do this by looking at the development of cities after they switch states. I take their analysis a step further by identifying special cases of switches: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct. Comparing the effects of these two types of switches allows me to tease out the effect of state power.

I make a key assumption: conquests imply state power differentials but successions do not. If state A takes a city from state B, 
absorbing a city by conquest is more likely if the new state is more powerful than the old one, but absorbing a city by succession underlies no such limitations - it only requires 
the new state after a conquest is selected for state power, but the new state after a succession is not. Roughly speaking, 
If a city is absorbed via conquest,

The dataset I use has several advantages. First, it traces the territorial history of cities in the Holy Roman Empire precisely up to a single year. Second, it contains

Rather than m

\cite{diamond1997} asserts that the geographical remoteness of China deprived it of the benefits of state competition that European states enjoyed throughout their history. \cite{landes2006} makes a similar argument, pointing out how the totalitarianism and bureaucratism that stifled innovation in medieval China would not have survived long in a more fragmented state system.

\cite{levine2013, levine2021} provide the theoretical foundation for conflict-driven selection, as well as compelling evidence for it. 
They argue that social evolution is driven by conflict. Their model finds that "evolution selects societies with high state power - the ability to prevail in conflict over rivals." (\citealp{levine2021}, p. 436).Their answer to Why Europe and not China or India: "Ghengis Khan, the cannon, and the English. Ghengis Khan accomplished two things: he brought cannons to Europe, and he depopulated Mongolia. In China the removal of the outside threat of the Mongols led - as the theory asserts - to an extractive hegemony. In Europe the advent of the cannon together with the preying of the English on the French and the Spanish resulted - as the theory asserts - in an inclusive balance of power." (p. 439). strong outsiders but primitive siege technology [cheap defensibility of cities] in India led to extractive balance of power.

@@

\cite{schoenholzer2022} use a dataset with a bigger geographical scope but lower resolution: they use urban population from \cite{bairoch1988} as a proxy for development, which is only available in intervals of 50 or 100 years, and a series of political maps.

@@

One paragraph about the history of Europe's state system; this is the place for Cervellati

One story about a city's territory switch

Two pithy figures, and a no-nonsense table

NO MORE THAN TWO PAGES TOTAL
